Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-07-08 paper/paper.tex
remove useless author info
replace list with set

\iffalse
%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended
\fi

@@

The structure of function @step@ is similar to the concrete CESK machine,
except the nondeterminism which makes @step@ return a list of reachable successor states.
We have two cases to consider (code shown below):

\begin{itemize}
  \item If the current control component is a @Let@, then the result of @App(f, ae)@ will
    be bound to variable @x@. In this case, we retrieve the set of closures that @f@ may
    represent. For each closure in the set, we perform nearly the same operations as in the
    concrete CESK machine, with an important difference: the continuation is allocated on
    the store @kstore@, so a new continuation address @new_kaddr@ must be constructed and
    a new frame @Frame(x, e, env)@ paired with the current continuation address @kaddr@ is
    merged into @new_kaddr@. Finally, a list of successor states is generated.

@@

\begin{lstlisting}
  def step(s: State): List[State] = {
    val new_time = s.tick
    s match {
      case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) =>
        val closures = atomicEval(f, env, bstore).toList
        for (Clos(Lam(v, body), env_c) <- closures) yield {
          val baddr = allocBind(v, new_time)
          val new_env = env_c + (v -> baddr)
          val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
          val new_kaddr = allocKont(body)
          val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
          State(body, new_env, new_bstore, new_kstore, new_kaddr, new_time)
        }
      case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
        val conts = kstore(kaddr).toList
        for (Cont(Frame(x, e, env_f), f_kaddr) <- conts) yield {
          val baddr = allocBind(x, new_time)
          val new_env = env_f + (x -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          State(e, new_env, new_store, kstore, f_kaddr, new_time)
        } } }
\end{lstlisting}

@@

\begin{lstlisting}
  def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
    case Nil => seen
    case hd::tl if seen.contains(hd) => drive(tl, seen)
    case hd::tl => drive(step(hd) ++ tl, seen + hd)
  }
  def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
\end{lstlisting}

@@

\begin{lstlisting}
  def step(s: State): List[State] = {
    val new_time = s.tick
    s match {
      case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
          val frame = Frame(x, e, env)
          val baddr = allocBind(v, new_time);  val new_env = env_c + (v -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          State(body, new_env, new_store, frame::konts, new_time)
        }
      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
        konts match {
          case Nil => List()
          case Frame(x, e, env_f)::konts =>
            val baddr = allocBind(x, new_time);  val new_env = env_f + (x -> baddr)
            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
            List(State(e, new_env, new_store, konts, new_time))
        } } }
\end{lstlisting}

>>>>

@@

The structure of function @step@ is similar to the concrete CESK machine,
except the nondeterminism which makes @step@ return a set of reachable successor states.
We have two cases to consider (code shown below):

\begin{itemize}
  \item If the current control component is a @Let@, then the result of @App(f, ae)@ will
    be bound to variable @x@. In this case, we retrieve the set of closures that @f@ may
    represent. For each closure in the set, we perform nearly the same operations as in the
    concrete CESK machine, with an important difference: the continuation is allocated on
    the store @kstore@, so a new continuation address @new_kaddr@ must be constructed and
    a new frame @Frame(x, e, env)@ paired with the current continuation address @kaddr@ is
    merged into @new_kaddr@. Finally, a set of successor states is generated.

@@

\begin{lstlisting}
  def step(s: State): Set[State] = {
    val new_time = s.tick
    s match {
      case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) =>
        val closures = atomicEval(f, env, bstore)
        for (Clos(Lam(v, body), env_c) <- closures) yield {
          val baddr = allocBind(v, new_time)
          val new_env = env_c + (v -> baddr)
          val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
          val new_kaddr = allocKont(body)
          val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
          State(body, new_env, new_bstore, new_kstore, new_kaddr, new_time)
        }
      case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
        for (Cont(Frame(x, e, env_f), f_kaddr) <- kstore(kaddr)) yield {
          val baddr = allocBind(x, new_time)
          val new_env = env_f + (x -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          State(e, new_env, new_store, kstore, f_kaddr, new_time)
        } } }
\end{lstlisting}

@@

\begin{lstlisting}
  def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
    case Nil => seen
    case hd::tl if seen.contains(hd) => drive(tl, seen)
    case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
  }
  def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
\end{lstlisting}

@@

\begin{lstlisting}
  def step(s: State): Set[State] = {
    val new_time = s.tick
    s match {
      case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore)) yield {
          val frame = Frame(x, e, env)
          val baddr = allocBind(v, new_time);  val new_env = env_c + (v -> baddr)
          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
          State(body, new_env, new_store, frame::konts, new_time)
        }
      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
        konts match {
          case Nil => Set()
          case Frame(x, e, env_f)::konts =>
            val baddr = allocBind(x, new_time);  val new_env = env_f + (x -> baddr)
            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
            Set(State(e, new_env, new_store, konts, new_time))
        } } }
\end{lstlisting}
